---
title: "Lab 1"
author: "An Bui"
date: "10 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Objectives:

- basic review (data wrangling)  
- intro to spatial data in R with `sf`  
- intro to Shiny apps  

#### 1. Attach packages
```{r set up packages, include = FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(tmap)
```

#### 2. Review: reading, exploring, warngling, visualizing data
##### a. Read in data.
```{r read in wine data, include = FALSE}
wine <- read_csv("wine_data.csv")
```

##### b. Wrangle data.
```{r wrangling, include = FALSE}
#### make new data set ####
# use original data set wine
wine_new <- wine %>%
  # select desired columns
  select(country, province, winery, region_1, points, price) %>% 
  # rename province column name to state
  rename(state = province) %>%
  # filter rows to only include California, Oregon, and Washington data
  filter(state == "California" |
           state == "Oregon" |
           state == "Washington") %>% 
  # add column for best price per points
  mutate(ppd = points/price) %>% 
  # arrange by decreasing points/dollar
  arrange(-ppd)
```

##### c. Use group_by() and summarize() to find summary statistics by group.
```{r summary statistics, echo = FALSE}
# make summary table from wine_new
wine_summary <- wine_new %>% 
  # create groups by state
  # btw you can group_by() for multiple columns
  group_by(state) %>% 
  # summary statistics, remove any NA values
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE)
  )

wine_summary
```

##### d. Visualize data.
```{r create plots, echo = FALSE}
#### histogram ####
# use ggplot, name data set, name count
wine_hist <- ggplot(wine_new, aes(x = ppd)) +
  # for any fill/point line where you want to base colors on variable, you have to start with aes()
  geom_histogram(aes(fill = state)) +
  # split graphs by state, set scales to change based on data distribution
  facet_wrap(~state, scales = "free")

wine_hist

#### scatterplot of price vs points with color determined by state, adding a linear model trendline on top ####
# make new plot using ggplot, use wine_new, name axes
wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  # make scatterplot, color points by state, change transparency to see points better
  # since transparency isn't attached to a variable, put it outside the aes()
  geom_point(aes(color = state), alpha = 0.5) +
  # add linear model on top, using linear model (ordinary least squares linear regression), without standard error
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  # change color scheme manually
  # scale_color_manual() (not the other ones) because you're setting a color (from line 84, in aes())
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  # separate graphs by state
  facet_wrap(~state) +
  # clean up plot background
  theme_light()

wine_points
```

#### 2. California counties spatial data using `sf`
All file types are important info within the spatial data (descriptions in key).
```{r spatial data files, include = FALSE}
#### call all files ####
# use st_read to read in files
ca_counties <- st_read(
  # set working directory (which you're already in)
  dsn = ".", 
  # call all files that start with CA_Counties_TIGER2016
  layer = "CA_Counties_TIGER2016")

#### select county, land columns ####
ca_land <- ca_counties %>% 
  select(NAME, ALAND)

#### add in population data ####
ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  # rename column to NAME to match ca_land
  rename(NAME = COUNTY)


```




























